### 2011-06-14 thesis/semantic-mapping.tex
More changes on the semantic mapping chapter... only introduction section missing now.

\begin{figure}[h]
\centering
\includegraphics[width=0.7\textwidth]{figures/dora-architecture.pdf}
\caption{\label{fig:spatial-knowledge}Layered structure of the spatial representation.
         The conceptual layer illustrates part of the ontology representing both instance
         and pre-defined world knowledge.}
\end{figure}

\subsubsection*{Sensory Layer}
The sensory layer keeps a detailed representation of the immediate surrounds of the agent.
That representation is based on directly sensed inputs as well on data-fusion for short
time\hyp{}intervals.

@@

\subsubsection*{Place Layer}
The place layer is responsible for maintaining a bottom\hyp{}up discretization of the
continuous space. On it the world is represented as a collection of discrete spatial
entities, called places, together with connectivity information.

@@

\subsubsection*{Categorical Layer}
The categorical layer holds the generic long-term, low-level representations of categorical models
of the agent's sensory information.
The knowledge represented by this layer is not instance specific. It is the knowledge
required to determine certain features indicate the presence of an entity of certain
category. E.g.\ the knowledge required to detect objects, landmarks or describe room appearances.
Other properties may as well be defined for example shape, color, edges and size properties
are all defined in terms of low\hyp{}level features.

@@

\subsubsection*{Conceptual Layer}
The conceptual layer provides an ontology that represents a taxonomy of the spatial concepts
and their relations with the properties represented on the categorical layer.
This associates semantic meaning to those properties that is useful for human-agent iteration
and provides relations that can be verbalized and explained with the ontology.

@@

\section{System Structure}
The \Gls{Dora} system~\cite{hanheide2011ijcai} consists of several co-operating
sub-systems, all of which actively use or maintain the spatial knowledge representation
(see \autoref{fig:dora-architecture}).

\subsection{Features}
\label{sec:dora-features}
The \emph{categorical layer} requires classifiers with the ability to categorize

\section{Conceptual Map}
\label{sec:conceptual-map}
As \gls{Dora} moves through the environment it builds a \emph{conceptual map}: a structural and
probabilistic representation of the space instantiated as a \emph{graphical model}.
It includes taxonomy of human\hyp{}compatible spatial concepts which are linked to the sensed 
instances of these concepts drawn from lower layers. It is the conceptual layer which 
contains the information that kitchens commonly contain cereal boxes and have certain 
general appearance and allows the robot to infer that the cornflakes box in front of the 
robot makes it more likely that the current room is a kitchen. The conceptual layer is 
described in terms of a probabilistic ontology defining spatial concepts and linking 
those concepts to instances of spatial entities (see the example of the ontology in
\autoref{fig:dora-architecture}).

>>>>

\begin{figure}[h]
\centering
\includegraphics[width=0.7\textwidth]{figures/spatial-knowledge-representation.pdf}
\caption{\label{fig:spatial-knowledge}Layered structure of the spatial representation.}
\end{figure}

\subsection{Sensory Layer}
The sensory layer keeps a detailed representation of the immediate surrounds of the agent.
That representation is based on directly sensed inputs as well on data-fusion for short
time\hyp{}intervals.

@@

\subsection{Place Layer}
The place layer is responsible for maintaining a bottom\hyp{}up discretization of the
continuous space. On it the world is represented as a collection of discrete spatial
entities, called places, together with connectivity information.

@@

\subsection{Categorical Layer}
The categorical layer holds the generic long-term, low-level representations of categorical models
of the agent's sensory information.
The knowledge represented by this layer is not instance specific. It is the knowledge
required to determine certain features indicate the presence of an entity of certain
category. E.g.\ the knowledge required to detect objects, landmarks or describe room appearances.
Other properties may as well be defined for example shape, color, edges and size properties
are all defined in terms of low\hyp{}level features.

@@

\subsection{Conceptual Layer}
The conceptual layer provides an ontology that represents a taxonomy of the spatial concepts
and their relations with the properties represented on the categorical layer.
This associates semantic meaning to those properties that is useful for human-agent iteration
and provides relations that can be verbalized and explained with the ontology.

@@

\section{System Structure}
The \Gls{Dora} system~\cite{hanheide2011ijcai} consists of several co-operating
sub-systems~(see \autoref{fig:dora-architecture}),
all of which actively use or maintain the \emph{spatial knowledge representation}.

The layer structure of the spatial knowledge representation allow to implement
data driven processes that control the flow of knowledge and information between the
lower and higher\hyp{}levels layers. To reduce computation and make it feasible not
all updates from lower\hyp{}levels are treated immediately. Instead the system waits
until a substantial change has occurred.

\begin{figure}[h]
\centering
\includegraphics[width=0.8\textwidth]{figures/dora-architecture.pdf}
\caption{\label{fig:dora-architecture}Dora the Explorer as well as the elements and the data flow inside
         the semantic mapping system.}
\end{figure}

Initially the low-level topology and metric mapping systems use a \gls{SLAM} algorithm~\cite{Folkesson07a}
that is used as representation on the sensory layer. Those two systems are also responsible
for creating the place map on the place layer, as that is highly related with the \gls{SLAM} goals.
By using the landmarks (doors) and other available information the system also tries to
group places and understand the concept of rooms.

At the same time the spatial property categorization system is responsible for detecting
instances of the categorical knowledge by running classifiers previously trained to detect
a set of features listed on \autoref{sec:dora-features}. For that it uses the data available
from a laser range finder and a camera and besides classifying the features it estimates also
its confidence on the information.
The extracted features together with the estimates is then integrated over time and space
in order to increase its robustness.

Finally the conceptual map system utilizes the acquired instance knowledge together with
the high\hyp{}level conceptual knowledge to create a probabilistic graphical model that allows
to run inferences.
Since this conceptual mapping processing is responsible for creating the probabilistic
graphical model it is described in detail in \autoref{sec:conceptual-map}.

\subsection{Features}
\label{sec:dora-features}
The \emph{categorical layer} requires classifiers with the ability to categorize the
low\hyp{}level features into categories.
For that \gls{Dora} performs detection on the following categorical entities:
\begin{description}
\item[Doorway Detection] by using the laser range finder the system tries to detect doors.
\item[Geometric Shape] by using features extracted from the laser data and pre-trained classifiers
the system tries to classify rooms according to their shapes (e.g.\ rectangular, square).
\item[Geometric Size] is classified once again recurring to laser data. In this case the system
\item[Object Detection] is performed by running traditional object recognition classifiers.
Those are previously trained on a supervised fashion to later be able to detect objects
based on their visual features, e.g.\ by using \gls{SIFT} as described in \autoref{sec:sift}.
tries to categorize the room size according to a set of previously learned concepts (e.g. large, medium, small).
\item[Visual Appearance] is extracted from the visual input. For that Dora uses a trained
a set of classifiers on global visual features such as \gls{CRFH} to categorize the visual
appearance of a room in categories such as corridor, office, library.
\end{description}

\section{Conceptual Map}
\label{sec:conceptual-map}
As \gls{Dora} moves through the environment it builds a \emph{conceptual map}: a structural and
probabilistic representation of the space instantiated as a \emph{graphical model}.
It includes taxonomy of human\hyp{}compatible spatial concepts which are linked to the sensed 
instances of these concepts drawn from lower layers. It is the conceptual layer which 
contains the information that kitchens commonly contain cereal boxes and have certain 
general appearance and allows the robot to infer that the cornflakes box in front of the 
robot makes it more likely that the current room is a kitchen. The conceptual layer is 
described in terms of a probabilistic ontology defining spatial concepts and linking 
those concepts to instances of spatial entities (see the example of the ontology in
\autoref{fig:spatial-knowledge}).
